---
title: "Lab 04 - Visualizing spatial data"
author: "Fiona Wang"
date: "02-13-2025"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
#install.packages("devtools")
#devtools::install_github("rstudio-education/dsbox")
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

Denny.   
There are `r nrow(dennys)` rows in the dataset and `r ncol(dennys)` variables in the dataset.
```{r Denny_properties}
head(dennys)
```

Each row corresponds to one restaurant. For each restaurant, the following variables were reported: address, city, state, zip, longitude, and latitude. 

### Exercise 2

La Quinta.  
There are `r nrow(laquinta)` rows and `r ncol(laquinta)` variables in the dataset.
```{r laquinta_properties}
head(laquinta)
```

Each row corresponds to one motel. Variables include: address, city, state, zip, longitude, and latitude.  

### Exercise 3

Looking from the website, there are La Quinta locations outside of the U.S. The countries that have La Quinta: Canada, Mexico, China, New Zealand, Turkey, United Arab Emirates, Chile, Colombia, and Ecuador. I can't tell from the webiste whether there are any Denny's outside of the U.S, because it only let me view U.S. locations. However, when I google, it says there are Denny's in Canada, Mexico, Puerto Rico, Phillipines, New Zealand, Honduras, UAE, and a couple more countries. 

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...

 
Add exercise headings as needed.

